Fast, Scalable Phrase-Based SMT Decoding

### 2016-05-09 acl.2016/acl2016.tex
changes

Therefore, we will focus on optimizing only the phrase-table and lexicalized reordering model.

@@

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use load-on-demand implementations of which Moses can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding time with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.

@@

\subsection{Lexicalized Reordering Model Optimizations}

Similar to the phrase-table, the lexicalized reordering model is trained on parallel data. A resultant model file is created in training which is then queried during decoding. Inevitably, the querying results in loss of decoding speed. Previous work such as~\newcite{junczys_tsd_2012b} improve querying speed with more efficient data structures.

However, the model's query keys are the source and target phrase of each translation rule. Rather than storing the lexicalized reordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file. This has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.

@@

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact data structure. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

@@

\subsection{Lexicalized Reordering Model}

The lexicalized reordering model requires a probability distribution of the reordering behaviour of each translation rule learnt from the training data. This is represented in the model file as a fixed number of probabilities for each rule, exactly how many probabilities is dependant on the model's parameterization during training. During decoding, a probability from this distribution is assigned to each hypothesis according to the reordering of the translation rule.

Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized reordering model distributions for each rule. During decoding, the probability distribution is then taken from the translation model instead of querying a separate file.

This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{lex-ro.png}} 
\end{tabular}
\caption{Decoding time with Compact Lexicalized Reordering, and integrated into a model the phrase-table}
\label{fig:lex-ro}
\end{figure} 
Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads. In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized reordering model\footnote{Strangely, using lexicalized reordering model results in slightly faster decoding than without the model. We double checked with other datasets. We suspect this may be an artefact of the algorithm or data structures, for example, more work is needed to resolve hypothesis recombination.}. Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores.

>>>>

Therefore, we will focus on optimizing only the phrase-table and lexicalized re-ordering model.

@@

For any phrase-table table of a realistic size, memory and loading time constraints requires us to use a load-on-demand implementation. Moses has several which we can make use of, each with differing performance characteristics. Figure~\ref{fig:moses-phrase-tables-time} shows the time taken to decode 800,000 sentences for the fastest two implementations.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-phrase-table.png}} 
\end{tabular}
\caption{Moses decoding time with two different phrase-table implementations}
\label{fig:moses-phrase-tables-time}
\end{figure} 
From this, it appears that the Probing phrase-table~\cite{Bogoychev:Thesis:2013} has the fastest translation rule lookup, especially with large number of cores, therefore, we will concentrate exclusively on this implementation from hereon.

@@

\subsection{Lexicalized Re-ordering Model Optimizations}

Similar to the phrase-table, the lexicalized re-ordering model is trained on parallel data. A resultant model file is created in training which is then queried during decoding. Inevitably, the querying results in loss of decoding speed. Previous work such as~\newcite{junczys_tsd_2012b} improve querying speed with more efficient data structures.

However, the model's query keys are the source and target phrase of each translation rule. Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file. This has precedent in ~\newcite{peitz2012jane} but the effect on decoding speed were not published. We will compare results with using a separate model in this paper.

@@

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All model files were then binarized; the language models were binarized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized re-ordering model using the compact data structure. These binary formats were chosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}

@@

\subsection{Lexicalized Re-ordering Model}

The lexicalized re-ordering model requires a probability distribution of the re-ordering behaviour of each translation rule learnt from the training data. This is represented in the model file as a fixed number of probabilities for each rule, exactly how many probabilities is dependant on the model's parameterization during training. During decoding, a probability from this distribution is assigned to each hypothesis according to the re-ordering of the translation rule.

Rather than holding the model probability distributions in the separate file, we pre-process the translation model file to include the lexicalized re-ordering model distributions for each rule. During decoding, the probability distribution is then taken from the translation model instead of querying a separate file.

This resulted in a significant decrease in decoding time, especially with high number of cores, Figure~\ref{fig:lex-ro}. 
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{lex-ro.png}} 
\end{tabular}
\caption{Decoding time with Compact Lexicalized Re-ordering, and integrated into a model the phrase-table}
\label{fig:lex-ro}
\end{figure} 
Decoding time decreased by 29\% with running with a single thread but it is over 5 times faster using 32 threads. In fact, the decoding time with the integrated model is similar to that \emph{without} a lexicalized re-ordering model\footnote{Strangely, using lexicalized re-ordering model results in slightly faster decoding than without the model. We double checked with other datasets. We suspect this may be an artefact of the algorithm or data structures, for example, more work is needed to resolve hypothesis recombination.}. Critically for systems with multicores servers, it enables the decoder to continue to scale, making efficient use of all available cores.
